Titeln på avhandlingens språk

### 2022-09-24 model.tex
notes
updating a bit

\subsection{Generating realistic execution times}\label{ssec:model:exectimes}

@@

We then assign a label to each sample indicating the class of the most recent transition in impairment level in the run.
In other words, for each recorded step, we look back until we find a step with a different level of impairment and assign a label to the current step according to whether that change was from a higher to a lower level of impairment or vice-versa.
However, if the current step is too far from the latest transition in impairment (by default, more than \num{8} steps), it is instead marked as \emph{no transition}; this label is also used for steps at the beginning of each run which have no previous transition in impairment.
This labeling allows us to capture the behavior described in \cref{item:remain} of \cref{ssec:plos}.

@@

Finally, in \cref{alg:timings} we present a step-by-step breakdown of how this data is subsequently used to generate realistic execution times.
In plain English, a model which uses this processed data to generate execution times needs to maintain variables for duration, previous impairment, and most recent transition.
At the beginning of each step, the model
\begin{enumerate}
    \item Checks whether any change in impairment has occurred.
          If it has, it resets duration to \num{1} and updates the transition variable accordingly.
          If not, it increases the duration variable by one unless that increase causes it to exceed the threshold of steps required to forget a transition.
          If that's the case, instead the model resets duration to \num{1} and sets the transition variable to ``no transition''.
    \item Discretizes the updated duration value.
    \item Finds all samples with binned neuroticism, impairment, duration, and transition values matching the current state of the system.
    \item Randomly samples the filtered data to obtain a realistic execution time value.
          \begin{itemize}
              \item Alternatively, a distribution can be first be fitted to the filtered samples.
                    The model then samples the theoretical distribution instead of the empirical samples.
          \end{itemize}
    \item Finally, updates the stored impairment value with the new impairment and proceeds to the next step.
\end{enumerate}

\begin{algorithm}
    \caption{Generating realistic execution times}\label{alg:timings}

@@

\begin{table*}
    \caption{Sample of the processed execution time data table.}\label{tab:data:exectime:processed}
    \begin{tabular}{rrrrcrccrc}
        \toprule
        {seq}        & {run\_id}    & {next\_exec\_time} & {ttf}        & {impairment} & {neuroticism\_raw} & {neuroticism} & {transition} & {duration\_raw} & {duration}   \\
        \midrule
        1            & 134146       & 3.655              & 0.000        & low          & 0.375              & low           & NoTransition & 1               & short        \\
        2            & 134146       & 4.439              & 0.597        & low          & 0.375              & low           & NoTransition & 2               & short        \\
        3            & 134146       & 2.943              & 0.554        & low          & 0.375              & low           & NoTransition & 3               & short        \\
        \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \)  & \( \cdots \) & \( \cdots \)    & \( \cdots \) \\
        169          & 134146       & 3.532              & 3.689        & high         & 0.375              & low           & Lower2Higher & 2               & short        \\
        1            & 134470       & 3.461              & 0.000        & low          & 0.594              & high          & NoTransition & 1               & short        \\
        2            & 134470       & 6.678              & 0.561        & low          & 0.594              & high          & NoTransition & 2               & short        \\
        \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \)  & \( \cdots \) & \( \cdots \)    & \( \cdots \) \\
        \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \)  & \( \cdots \) & \( \cdots \)    & \( \cdots \) \\
        169          & 137353       & 4.615              & 0.579        & low          & 0.625              & high          & NoTransition & 2               & short        \\
        \bottomrule
    \end{tabular}
\end{table*}

\begin{figure*}
    \centering%
    % \hspace{\fill}%
    \begin{subfigure}[t]{.28\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./model_data/imp_neur_vs_exectime.png}
        \caption{}\label{fig:timing:impneurvsetime}
    \end{subfigure}%
    \hspace{\fill}%
    \begin{subfigure}[t]{.28\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./model_data/duration_vs_exectime.png}
        \caption{}\label{fig:timing:durvsetime}
    \end{subfigure}%
    \hspace{\fill}%
    \begin{subfigure}[t]{.28\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./model_data/impairment_transition_vs_exectime.png}
        \caption{}\label{fig:timing:imptransvsetime}
    \end{subfigure}%
    % \hspace{\fill}%
    \caption{%
        Visualizations of the model data after discretization and post-processing.
        \subref{fig:timing:impneurvsetime} illustrates the effect of impairment on execution times, and how neuroticism modulates this behavior;
        \subref{fig:timing:durvsetime} exemplifies how the model captures the speed-up and slow-down behaviors due to impairment and duration described in \textcite{olguinmunoz:impact2021};
        finally, \subref{fig:timing:imptransvsetime} shows the lingering effects of impairment after a transition.
        Error bars indicate the \SI{95}{\percent} \ac{CI}.
    }\label{fig:timing}
\end{figure*}

@@

\subsection{Model verification}\label{ssec:model:verification}
\todo[inline]{Needs better name.}

>>>>

\subsection{Generating realistic execution times}\label{ssec:model:exectimes}
\todo[inline]{Rework with new tweaked model.}

@@

We then assign a label to each sample indicating the class of the most recent transition in impairment level in the run.
In other words, for each recorded step, we look back until we find a step with a different level of impairment and assign a label to the current step according to whether that change was from a higher to a lower level of impairment or vice-versa.
% However, if the current step is too far from the latest transition in impairment (by default, more than \num{8} steps), it is instead marked as \emph{no transition}; this label is also used for steps at the beginning of each run which have no previous transition in impairment.
% This labeling allows us to capture the behavior described in \cref{item:remain} of \cref{ssec:plos}.

@@

Finally, in \cref{alg:timings} we present a step-by-step breakdown of how this data is subsequently used to generate realistic execution times.
In plain English, a model which uses this processed data to generate execution times needs to maintain variables for duration, previous impairment, and most recent transition.
At the beginning of each step, the model
\begin{enumerate}
    \item Checks whether any change in impairment has occurred.
          If it has, it resets duration to \num{1} and updates the transition variable accordingly.
          If not, it increases the duration variable by one.
        %   unless that increase causes it to exceed the threshold of steps required to forget a transition.
        %   If that's the case, instead the model resets duration to \num{1} and sets the transition variable to ``no transition''.
    \item Discretizes the updated duration value.
    \item Finds all samples with binned neuroticism, impairment and duration matching the state of the system.
    If, and only if, the current duration falls within the \emph{short} duration bin, this data is further bisected according to the current value of the transition variable.
    \item Randomly samples the filtered data to obtain a realistic execution time value.
          \begin{itemize}
              \item Alternatively, a distribution can be first be fitted to the filtered samples.
                    The model then samples the theoretical distribution instead of the empirical samples.
          \end{itemize}
    \item Finally, updates the stored impairment value with the new impairment and proceeds to the next step.
\end{enumerate}

\begin{algorithm}
    \todo[inline]{Update algorithm.}
    \caption{Generating realistic execution times}\label{alg:timings}

@@

\begin{table*}
    \todo[inline]{Remove this table, or re-do it.}
    \caption{Sample of the processed execution time data table.}\label{tab:data:exectime:processed}
    \begin{tabular}{rrrrcrccrc}
        \toprule
        {seq}        & {run\_id}    & {next\_exec\_time} & {ttf}        & {impairment} & {neuroticism\_raw} & {neuroticism} & {transition} & {duration\_raw} & {duration}   \\
        \midrule
        1            & 134146       & 3.655              & 0.000        & low          & 0.375              & low           & NoTransition & 1               & short        \\
        2            & 134146       & 4.439              & 0.597        & low          & 0.375              & low           & NoTransition & 2               & short        \\
        3            & 134146       & 2.943              & 0.554        & low          & 0.375              & low           & NoTransition & 3               & short        \\
        \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \)  & \( \cdots \) & \( \cdots \)    & \( \cdots \) \\
        169          & 134146       & 3.532              & 3.689        & high         & 0.375              & low           & Lower2Higher & 2               & short        \\
        1            & 134470       & 3.461              & 0.000        & low          & 0.594              & high          & NoTransition & 1               & short        \\
        2            & 134470       & 6.678              & 0.561        & low          & 0.594              & high          & NoTransition & 2               & short        \\
        \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \)  & \( \cdots \) & \( \cdots \)    & \( \cdots \) \\
        \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \) & \( \cdots \) & \( \cdots \)       & \( \cdots \)  & \( \cdots \) & \( \cdots \)    & \( \cdots \) \\
        169          & 137353       & 4.615              & 0.579        & low          & 0.625              & high          & NoTransition & 2               & short        \\
        \bottomrule
    \end{tabular}
\end{table*}

\begin{figure*}
    \centering%
    % \hspace{\fill}%
    \begin{subfigure}[t]{.28\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./model_data/imp_neur_vs_exectime.png}
        \caption{}\label{fig:timing:impneurvsetime}
    \end{subfigure}%
    \hspace{\fill}%
    \begin{subfigure}[t]{.28\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./model_data/duration_vs_exectime.png}
        \caption{}\label{fig:timing:durvsetime}
    \end{subfigure}%
    \hspace{\fill}%
    \begin{subfigure}[t]{.28\textwidth}
        \centering
        \includegraphics[width=\textwidth]{./model_data/impairment_transition_vs_exectime.png}
        \caption{}\label{fig:timing:imptransvsetime}
    \end{subfigure}%
    % \hspace{\fill}%
    \caption{%
        Visualizations of the model data after discretization and post-processing.
        \subref{fig:timing:impneurvsetime} illustrates the effect of impairment on execution times, and how neuroticism modulates this behavior;
        \subref{fig:timing:durvsetime} exemplifies how the model captures the speed-up and slow-down behaviors due to impairment and duration described in \textcite{olguinmunoz:impact2021};
        finally, \subref{fig:timing:imptransvsetime} shows the lingering effects of impairment after a transition, by comparing the mean execution times of the steps in the first bin of duration after a transition.
        Error bars indicate the \SI{95}{\percent} \ac{CI}.
    }\label{fig:timing}
\end{figure*}

@@

\subsection{Model verification}\label{ssec:model:verification}
\todo[inline]{Needs better name.}
\todo[inline]{Rework with new tweaked models.}
